Dismal Code: Studying the Evolution of Security Bugs

### 2013-02-06 docs/msr.tex
work on the inital results subsection

\title{Dismal Code: Studying the Evolution of Security Vulnerabilities}

@@

\begin{abstract}
%\boldmath
Security bugs are critical programming errors that can lead to serious
vulnerabilities in software. Such bugs may allow an attacker to take over
an application, steal data or prevent the application from working at all.
In this paper, we present how we used the Maven repository to study the
characteristics of security bugs individually and in relation to other software
bugs. In particular, we analyzed every artifact of the repository by using a
popular static analysis tool. Then by taking advantage of the various features of
the repository (multiple versions for every project, dependencies) we studied
the evolution of security bugs through time and their relationship with the
size of the corresponding artifact and other categories. Also, based on the
dependencies of artifacts we constructed the graph of the repository and
studied the most popular nodes in terms of their PageRank. 
\end{abstract}

\begin{IEEEkeywords}
Software Vulnerabilities, Static Analysis, Software Evolution, Software
Security, Maven, FindBugs.
\end{IEEEkeywords}

@@

One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it. For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a popular static analysis tool that has already been used in
research~\cite{AP10, HP07, HP04}. Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the Maven\footnote{\url{http://maven.apache.org/}}
repository (approximately 265GB of data). Then we observed the changes that
involved the security bugs and their characteristics. This research builds upon
our earlier work on the topic~\cite{MGS12}.

@@

First, we scanned the Maven repository for appropriate {\sc jar}s.
We discuss the {\sc jar} selection process in the next 
section. With this list we created a series of processing tasks
and added them to the task queue. Then we
executed 25 (Unix-based) workers that checked out tasks from the queue,
processed the data and stored the results to the data repository.

@@

In addition, we implemented a series of checks in the worker scripts that
checked if the {\sc jar}s are valid in terms of implementation. For instance,
for every {\sc jar} the worker checked if there were any {\it .class} files
before invoking FindBugs. After the project filtering we narrowed down
our data set to 17,055 projects with 114,399 versions.
Table \ref{tbl:repository} summarises the data set information and
provides the basic descriptive statistic measurements. The distribution of version
count among the selected projects is presented in Figure \ref{fig:version-count}.

@@

The repository has many project with a
few number of versions. On the contrary, there are numerous projects with a
number of versions that is less than ten.
The statistical mesurements presented in Table \ref{tbl:repository}
indicate that we have 17,505 projects total and the data sets median
is 3, which means that almost 50\% (8,753 projects) of the project
population has 1 to 3 versions. There are a few projects containing ten
versions and only a few with hundreds of versions. The maximum number of
versions for a project is 337. The 3$^{rd}$ quartile measurement
also indicated that 75\% (13,129) of the projects have a maximum of 8 versions.

@@

@@

\begin{figure*}
	\centering
	\includegraphics[scale=0.6]{bug_percent}
	\caption{Bug Percentage in Maven Repository}
	\label{fig:bug-percentage}
\end{figure*}

\begin{table}
\centering
\caption{Bug categorisation according to FindBugs}
\label{tbl:bug-categorisation}
\begin{tabular}{l p{15em}}
\hline
Category & Description\\
\hline
Bad Practice & Violations of recommended and essential
coding practice. \\
Correctness & Involves coding mistakes resulting in code
that was probably not what the developer intended. \\
Experimental & Includes unsatisfied obligations. For instance,
forgetting to close a file \\
Internationalisation & Indicates the use of non-localized methods \\
Multi-Threaded Correctness & Thread synchronization issues. \\
Performance & Involves inefficient memory usageallocation, usage 
of non-static classes \\
Style & Code that is confusing, or
written in a way that leads to errors.\\
Security & Contains two subcategories, namely: {\it Security} and {\it
Malicious Code}. Involves code that could lead to security breach. \\
\hline
\end{tabular}
\end{table}

@@

{\it Querying techniques} are used to answer a broad range of questions
regarding the evolution history of a project~\cite{HG05}. Bhattacharya et
al.~\cite{BN11, B11} proposed a framework that is based upon
recursively enumerable languages. The framework can correlate software
bugs with developers in various ways. For instanse, return the list of
bugs fixed by a specific developer. Fischer et al.~\cite{FPG03} proposed
an approach for populating a release history database that combines code
information with bug tracking data. In this way, a developer can couple files
that contain common bugs, estimate code maturity with respect to the bugs
and others. The ``Ultimate Debian Database''~\cite{NZ10} is an {\sc sql}-based
framework that integrates information about the Debian project from various
sources to answer queries related to software bugs and source code.

>>>>

\title{Dismal Code: Studying the Evolution of Security Bugs}

@@

\begin{abstract}
%\boldmath
Security bugs are critical programming errors that can lead to serious
vulnerabilities in software. Such bugs may allow an attacker to take over
an application, steal data or prevent the application from working at all.
In this paper, we present how we used the Maven repository to study the
characteristics of security bugs individually and in relation to other software
bugs. In particular, we analyzed every artifact of the repository by using a
popular static analysis tool. Then by taking advantage of the various features of
the repository (multiple versions for every project, dependencies and others) we studied
the evolution of security bugs through time and their relationship with a) the
size of the corresponding artifact and b) other bug categories. Also, based on the
dependencies of artifacts we constructed the graph of the repository and
examined the most popular nodes based on their PageRank. 
\end{abstract}

\begin{IEEEkeywords}
Software Bugs, Static Analysis, Software Evolution, Software
Security, Maven, FindBugs.
\end{IEEEkeywords}

@@

One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it. For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10, HP07, HP04}. Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the Maven\footnote{\url{http://maven.apache.org/}}
repository (approximately 265GB of data). Then we observed the changes that
involved the security bugs and their characteristics. This research builds upon
our earlier work on the topic~\cite{MGS12}.

@@

First, we scanned the Maven repository for appropriate {\sc jar}s.
We discuss the {\sc jar} selection process in the next 
section. With this list we created a series of processing tasks
and added them to the task queue. Then we
executed twenty five (Unix-based) workers that checked out tasks from the queue,
processed the data and stored the results to the data repository.

@@

In addition, we implemented a series of checks in the worker scripts that
checked if the {\sc jar}s are valid in terms of implementation. For instance,
for every {\sc jar} the worker checked if there were any {\it .class} files
before invoking FindBugs. After the project filtering we narrowed down
our data set to 17,055 projects with 114,399 versions.
Table~\ref{tbl:repository} summarises the data set information and
provides the basic descriptive statistic measurements. The distribution of version
count among the selected projects is presented in Figure \ref{fig:version-count}.

@@

The repository has many project with a
few number of versions. On the contrary, there are numerous projects with a
number of versions that is less than ten.
The statistical mesurements presented in Table~\ref{tbl:repository}
indicate that we have 17,505 projects total and the data sets median
is 3, which means that almost 50\% (8,753 projects) of the project
population has 1 to 3 versions. There are a few projects containing ten
versions and only a few with hundreds of versions. The maximum number of
versions for a project is 337. The 3$^{rd}$ quartile measurement
also indicated that 75\% (13,129) of the projects have a maximum of 8 versions.

@@

Our findings can be analyzed at two levels. First, we discuss some
primary observations concerning the security bugs of the Maven repository as a whole.
Then, we provide a comprehensive analysis of the results and highlight our key findings.

@@

FindBugs separates software bugs into seven categories (see
Table~\ref{tbl:bug-cat}). Two of them involve security: {\it Security} and {\it
Malicious Code}. Our early findings indicated that from the total number of the
artifacts, 4,270 of them contained at least one bug coming from the first category
and 45,145 from the second.

\begin{table}
\centering
\caption{Bug categorisation according to FindBugs}
\label{tbl:bug-cat}
\begin{tabular}{l p{15em}}
\hline
Category & Description\\
\hline
Bad Practice & Violations of recommended and essential
coding practice. \\
Correctness & Involves coding mistakes resulting in code
that was probably not what the developer intended. \\
Experimental & Includes unsatisfied obligations. For instance,
forgetting to close a file \\
Internationalisation & Indicates the use of non-localized methods \\
Multi-Threaded Correctness & Thread synchronization issues. \\
Performance & Involves inefficient memory usageallocation, usage 
of non-static classes \\
Style & Code that is confusing, or
written in a way that leads to errors.\\
Security & Contains two subcategories, namely: {\it Security} and {\it
Malicious Code}. Involves code that could lead to security breach. \\
\hline
\end{tabular}
\end{table}\begin{figure*}
	\centering
	\includegraphics[scale=0.6]{bug_percent}
	\caption{Bug Percentage in Maven Repository}
	\label{fig:bug-per}
\end{figure*}

Figure~\ref{fig:bug-per} shows how software bugs are distributed among the
repository. Together with the {\it Bad Practice} bugs and the {\it Style} bugs,
security bugs (the sum of the {\it Security} and {\it Malicious Code}
categories) are the most popular in the repository.

\begin{figure*}
	\centering
	\includegraphics[scale=0.6]{bug_percent}
	\caption{Bug Percentage in Maven Repository}
	\label{fig:bug-percentage}4,270 
\end{figure*}

@@

{\it findbugs Querying techniques} are used to answer a broad range of questions
regarding the evolution history of a project~\cite{HG05}. Bhattacharya et
al.~\cite{BN11, B11} proposed a framework that is based upon
recursively enumerable languages. The framework can correlate software
bugs with developers in various ways. For instanse, return the list of
bugs fixed by a specific developer. Fischer et al.~\cite{FPG03} proposed
an approach for populating a release history database that combines code
information with bug tracking data. In this way, a developer can couple files
that contain common bugs, estimate code maturity with respect to the bugs
and others. The ``Ultimate Debian Database''~\cite{NZ10} is an {\sc sql}-based
framework that integrates information about the Debian project from various
sources to answer queries related to software bugs and source code.
